---
title: "dataLabelWL"
output: html_notebook
---

```{r}
library(dplyr)
library(readr)
library(lubridate)

# Setting the dir path
setwd("/Users/abyvelazquez/Developer/Repos/DS_TennisSet/data")
data_M <- read.csv("frenchopenFinalMTransformed.csv")
```
Convert the Date attribute into a Date Object and not character, also order data and limit to 2015 to now window.
```{r}
data_M$Date <- dmy(data_M$Date)
# Sort by Date of match then by round
data_M[
  with(data_M, order(data_M$Date, data_M$Round)),
]
DelimiterDate <- as.Date("2016/01/01")
data_F <- data_M[data_M$Date >= DelimiterDate ,]


```
Now we want to create a label (1->won, 0-> lost) that indicates whether the winner in fact won based on a rule.
The rule can begin to be very simple, e.g: if Wrank is smaller than LRank, its better positioned in the begining of
the tournament, therefore it should win.
```{r}
data_F$LabelWinner <- ifelse((data_F$WRank - data_F$LRank < 0), 1, 0)
data_F
```

Our goal is to create a more complex rule, taking into account not only the Rankings of each player, but other
attributes such as the Pts they have and the odds.
Later we want to take into account former matches of the players and even the number of rounds they played in each
tournament indicating how far into the finals they get.
