---
title: "dataLabelWL"
output: html_notebook
---

```{r}
library(dplyr)
library(readr)
library(lubridate)

# Setting the dir path
setwd("/Users/abyvelazquez/Developer/Repos/DS_TennisSet/Roland-Garros/data")
data_M <- read.csv("frenchopenFinalMTransformed.csv")
```
Convert the Date attribute into a Date Object and not character, also order data and limit to 2015 to now window.
```{r}
data_M$Date <- dmy(data_M$Date)
# Sort by Date of match then by round
data_M[
  with(data_M, order(data_M$Date, data_M$Round)),
]
DelimiterDate <- as.Date("2016/01/01")
data_F <- data_M[data_M$Date >= DelimiterDate ,]


```
Now we want to create a label (1->won, 0-> lost) that indicates whether the winner in fact won based on a rule.
The rule can begin to be very simple, e.g: if Wrank is smaller than LRank, its better positioned in the begining of
the tournament, therefore it should win.
```{r}
data_F$LabelWinnerRank <- ifelse((data_F$WRank - data_F$LRank < 0), 1, 0)
data_F
```

Our goal is to create a more complex rule, taking into account not only the Rankings of each player, but other
attributes such as the Pts they have and the odds.
Later we want to take into account former matches of the players and even the number of rounds they played in each
tournament indicating how far into the finals they get.
```{r}
# To get the frequency of our label, 254 Ceros and 633 Ones:
table(data_F$LabelWinnerRank)
```
```{r}
data_F$LabelWinnerPoints <- ifelse((data_F$WPts > data_F$LPts), 1, 0)
data_F
```
```{r}
# To get the frequency of our label, 254 Ceros and 633 Ones:
table(data_F$LabelWinnerPoints)
```
Interesting... we found both rules have the same frequency, is it equal for every row?
```{r}
data_F$SameLabelWPWR <- ifelse((data_F$LabelWinnerPoints == data_F$LabelWinnerRank), 1, 0)
table(data_F$SameLabelWPWR)
```
We can see that if a player has a better Rank, it also means that he always has a higher number of points than
his opponent.
Therefore, I dont see the case of combining a rule that takes into account Rank and Points into one Label of W/L.




Lets now see if we create a simple rule taking into account the odds.
First we want to eliminate NA rows where the odds attributes are na, so it does not interfere with our logic of the rule
```{r}
data_F[!is.na(data_F$B365W),]
data_F[!is.na(data_F$B365L),]
data_F[!is.na(data_F$PSW),]
data_F[!is.na(data_F$PSL),]
```
For the B365 odds label, because we know it is using the Decimal European odds system, we know that the favorite player
will always have a minor value, and the underdog player(the one where fans and experts say is the less likely to win,
will have a mayor turnover of investment in the likelihood he wins)
```{r}
data_F$LabelWinnerOddsB65 <- ifelse((data_F$B365W < data_F$B365L), 1, 0)
# We can see that taking into account B365 odds, its a better rule than ranking and points alone.
# With 210 Ceros and 679 Ones.
table(data_F$LabelWinnerOddsB65)
```
Lets try a label with the the PS odds
```{r}
data_F$LabelWinnerOddsPS <- ifelse((data_F$PSW < data_F$PSL), 1, 0)
# We can see that taking into account PS odds, its a better rule than ranking and points alone.
# With 202 Ceros and 687 Ones.
table(data_F$LabelWinnerOddsPS)

```
Lets try with MaxW and MaxL odds.
```{r}
data_F$LabelWinnerOddsMAX <- ifelse((data_F$MaxW < data_F$MaxL), 1, 0)
# We can see that taking into account Max Winner odds, its a better rule than ranking and points alone.
# With 201 Ceros and 687 Ones.
table(data_F$LabelWinnerOddsMAX)
```
Lets try with AvgW and AvgL odds
```{r}
data_F<-data_F[!is.na(data_F$AvgW),]
data_F<-data_F[!is.na(data_F$AvgL),]
data_F$LabelWinnerOddsAvg <- ifelse((data_F$AvgW < data_F$AvgL), 1, 0)
# We can see that taking into account Avg Winner odds, its a better rule than ranking and points alone.
# With 202 Ceros and 686 Ones.
table(data_F$LabelWinnerOddsAvg)
```
Create a new csv with the simple labels we want to try
```{r}

# Convert Date to character again for promidat processing, the date as d-%m-%Y, causes error loading the data.
data_F$Date <- as.character(data_F$Date,"%d/%m/%Y" )
data_F_simple_labels <- data_F[,!names(data_F) %in% c("SameLabelWPWR", "PromDif" )]
write.csv(data_F_simple_labels, "data_simple_labels.csv", row.names=FALSE)
```
Another way to create our own rank and then include it in the rule would be to create a label that accounts for the
number of times the player has won in that specific round he is playing, e.g:
If Juan Perez is playing the 1st round, how many times has he won in the past? 3 times
If Ronaldo Sanchez is playing the 1st round against JP, how many times has he won in the past? 5 times
In this case, the label 1 would go to player Ronaldo because he has won the most 1st rounds in past tournaments.

```{r}
# There are several ways to calculate that value
# One way would be to filter the first rounds in the dataset
# go row by row
# Start a counter from 0 at the beginning for each player
# When the name of a player is the same as the one we are evaluating in "Winner" attribute, then:
# add 1 to the counter and put it as a value in a new table
# With id date, attribute Name of player, attribute named as "Won1stRounds"
# Then for each match, we do a search for the latest date but before the start of that match for both players
# The player that has the most wins for that round is labeled as 1.
# HELP WITH THIS.
```
